# Auto-generating Paint-by-Numbers Instructions {#paintByNumbers}

## Introduction

This project aims to convert an image (e.g., of my cat) into paint-by-number instructions. This involves:

* reducing the number of colors down to a reasonable number, i.e. reducing the 16+ million colors possible with 8-bit images down to like 10-15,

* converting raster data into polygons based on those limited colors, and

* saving those resulting polygons with their color IDs as printable and readable instructions.

This document follows the adventure of getting it all to work.

## Packages being used

The packages we load are all on CRAN. `raster` has handy functions for handling and viewing images, `sf` and `smoothr` help with handling the polygons, and `tidyverse` has general grammar for tidy data manipulation. Disclaimer: there are probably more packages out there that do similar things, but I'm coming from a geospatial day job and I'm using these packages since there are already familiar to me. Other packages needed for dependencies are `rgdal` and `rgeos`.

```{r message=FALSE}
library(raster)
library(sf)
library(tidyverse)
library(smoothr)
library(units)
library(ggthemes)  # add to text above
library(wesanderson)
library(rdist)
```

## Step 1: Reading in the image

```{r }
img <- brick("../images/byxbee_loaf.jpg") %>% 
  aggregate(25) # reducing the resolution up front to save time
plotRGB(img)
```

## Step 2: Reducing the number of colors

### Attempt 1: just rounding the individual RGB values

```{r}
cutoff <- 75  # number to rount to
rounded_img <- img  # make a copy of the orignal image object

values(rounded_img) <- values(img)  %>%  # extract values from image (strip spatial component)
  as_tibble() %>% #converting the matrix to tibble for tidy data manipulation
  mutate_all(~cutoff*floor(./cutoff)) %>% # overwrite all values to rounded down values to cutoff
  as.matrix() # convert back to matrix

plotRGB(rounded_img)
```

Well it worked in that we can still see the cat and there are <20 unique colors, but its not the ideal result. 

### Attempt 2: using kmeans clustering 
```{r message=FALSE, warning=FALSE}
N <- 8 # number of colors
img_clust <- values(img) %>%
  kmeans(N, iter.max = 50)
centroids_img <- img
values(centroids_img) <- img_clust$cluster  %>%
  enframe(name = NULL) %>%
  left_join(img_clust$centers %>% 
              as_tibble(rownames = "value") %>%
              mutate(value = as.numeric(value)))  %>%
  mutate_all(~round(.,0)) %>%
  as.matrix()

plotRGB(centroids_img[[-1]]) # I'm skipping the first layer that contains the cluster ID
```

That's a bit better. But what if we have a set palette we want to use? 

### Attempt 3: specifying a palette first

I haven't tried this yet, but the plan is to see if we can split the colors from the image into a pre-defined set of colors. Maybe cool ones like the [Wes Anderson pallete](https://github.com/karthik/wesanderson) or matching to paint sets to make painting the printed result more feasible. 

An initial attempt with the Isle of Dogs palette:

```{r, message=FALSE}
N <- 15  # number of colors
(wes_pal <- wes_palette("IsleofDogs1", N, type = "continuous"))

wes_rgb <- wes_pal[1:N] %>%
  col2rgb() %>%
  t() 

img_wes <- values(img) %>%
  cdist(wes_rgb) %>%
  as_tibble() %>%
  rowwise() %>%
  mutate(cluster = which.min(c_across(everything()))) %>%
  pull(cluster)

def_pal_img <- img
values(def_pal_img) <- img_wes  %>%
  enframe(name = NULL) %>%
  left_join(wes_rgb %>% 
              as_tibble(rownames = "value") %>%
              mutate(value = as.numeric(value)))  %>%
  mutate_all(~round(.,0)) %>%
  as.matrix()

plotRGB(def_pal_img[[-1]])

```



## Step 3: Converting the image into polygons of the same color
```{r message=FALSE}
poly_img <- rasterToPolygons(def_pal_img$value,
                             dissolve = TRUE,
                             digits = 0)
```

```{r, message=FALSE, warning=FALSE}
st_img <- poly_img %>%
  st_as_sf() %>%
  st_buffer(5) %>%
  st_simplify(TRUE,dTolerance = 15) %>% #simplifies lines by removing vertices
  #drop_crumbs(threshold = 10 ) %>% # these sound applicable but I keep getting error messages!
  #fill_holes(threshold = 10) %>%
  smooth(method = "ksmooth", smoothness = 0.5) %>%
  st_cast("POLYGON")

# clust_cols <- img_clust$centers %>% 
#   as_tibble(rownames = "value") %>%
#   mutate(hex = rgb(byxbee_loaf.1/255,byxbee_loaf.2/255,byxbee_loaf.3/255)) %>%
#   pull(hex)
# 
# img_pal <- function(n){
#   return(clust_cols[1:n])
# }

# Try to visualize what the painting would look like
st_img %>%
  plot(pal = wes_pal[1:N],breaks=0:15,  # use the palette from the previous section
       border = NA,  # no polygon outlines 
       key.pos=NULL, # no legend
       main = NULL)  # no title
```

So now we need to print the instructions and make sure they are readable on a standard printer paper to make things easy for people. I'm switching to `ggplot` plotting because it's easier to customize visualizations.
```{r}
st_img %>%
  ggplot() +
  geom_sf(fill = NA,
       color = "black", # draw lines for polygons
       size = 0.1) + # fine lines
  geom_sf_text(aes(label = value), # label color ID per polygon
               size = 1) +  
  theme_map()  # no axes or anything like that
```

This seems like a lot of polygons! We will have to iteratively tweak our simplifications.

## Future expansions:

 * Consider using a machine learning classification approach?

 * Have an option to split skinny images (e.g. panoramic photos) into multiple pages of instructions, which would be helpful for breaking up mural instructions
 
 * Have an option to have a grid included on the instructions in case people need to transfer the outlines to a large area, e.g. a wall for a mural
 
 * Have an option to fit a custom piece of paper